---
title: "Homework 4"
author: "Kunaal Raghav"
format: pdf
editor: visual
---

## Install and Library Packages

```{r}
#install.packages("pROC")
library(pROC)
library(ISLR2)
library(MASS)
```

## Stock Market Prediction (Exercise 4.8.13)

Question A

```{r}
data("Weekly")

Directions <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5,
                data = Weekly, family = binomial)
summary(Directions)
```

Right now the only variable that seems to be statistically significant is Lag2, or the percentage returns for the previous two weeks.

Part B

```{r}
probs <- predict(Directions, type = "response")
preds <- ifelse(probs > 0.5, "Up", "Down")
table(preds, Weekly$Direction)
mean(preds == Weekly$Direction)
```

On one hand, there are a lot of false negatives, meaning that predictions are unusually high for up/ raises in the markets, where as they actually come to be heavily in the downs/ negative.

Part C

```{r}
train <- Weekly$Year <= 2008
test  <- Weekly$Year > 2008

fit_lag2 <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)
probs_test <- predict(fit_lag2, newdata = Weekly[test, ], type = "response")
preds_test <- ifelse(probs_test > 0.5, "Up", "Down")

table(preds_test, Weekly$Direction[test])
mean(preds_test == Weekly$Direction[test])
```

Considering that this model has a 62.5%, we can say that this model is more accurate. With the extra variables that were removed, its safe to say that there is a pattern of over fitting in the previous model

Part D

```{r}

# Train/test split
train <- Weekly$Year <= 2008
test  <- Weekly$Year > 2008
fit_lag2 <- glm(Direction ~ Lag2, data = Weekly,
                family = binomial, subset = train)
probs_test <- predict(fit_lag2, newdata = Weekly[test, ], type = "response")
actual_test <- Weekly$Direction[test]
roc_obj <- roc(actual_test, probs_test, levels = c("Down", "Up"))
plot(roc_obj, col = "blue", lwd = 2,
     main = "ROC Curve for Logistic Regression (Lag2 Only, Test Data)")
abline(0, 1, lty = 2, col = "gray")

```

The model, while slightly better than a random prediction, which sits at .5, is not well established enough to properly predict down markets.

## LDA vs QDA (Exercise 4.8.5)

Part A

On the training data set, since the model is more flexible, QDA is better used

On the test data set, since the model matches that classifer, LDA is better , as it will produce a lower test error

Part B

In the training set, QDA will have a lower error rate, because it will fit the model better than LDA, due to flexibility

With a nonlinear setting in the Bayes decision boundary, QDA will preform better due to low bias

Part C

With increases in the n size, test performance of QDA should improve, as QDA has high variance. Small n sizes means the variance hurts test performance

Part D

The aforementioned statement is wrong, as , the Bayes Classifier, in its Linear form, matches LDAS which has low bias and lower variance ( as compared to QDA).

## Non Uniform Prior (Exercise 4.8.7)

The following is the probability that we are setting up for calculation

$$
P (Dividend | X=4)
$$

$$
\mu_1=10
$$

$$
\mu_2=0
$$

$$
\sigma=6
$$

$$
\sigma^2=36
$$

Prior Probability issuing a dividend:

$$
P(D)=.8
$$

Prior Probability not issuing a dividend:

$$
P(N)=.2
$$

Observed Profit

$$
X=8
$$

The first step is calculating for the following equations:

$$
f_1(4)=f(X=4|D)
$$

$$
f_1(4)=f(X=4|N)
$$

We use the density function to calculate. The density function is:

$$
f(x)=\frac{1}{\sigma\sqrt{2\pi}}^{(-\frac{(x-\mu)^2}{2\sigma^2})}
$$

For dividend companies, the filled in equation is as follows:

$$
f_2(4)=\frac{1}{6\sqrt{2\pi}}^{(-\frac{(4-10)^2}{72})}
$$

The transformed equation becomes:

$$
f_1(4)=\frac{1}{6\sqrt{2\pi}}e^{-.5}
$$

For non-dividend companies, the filled in equation is as follows

$$
f_2(4)=\frac{1}{6\sqrt{2\pi}}^{(-\frac{(4-0)^2}{72})}
$$

The transformed equation becomes:

$$
f_2(4)=\frac{1}{6\sqrt{2\pi}}e^{-.22}
$$

Then we plug in the Bayes Theorem:

$$
P(D|X=4)= \frac {f_1(4)P(D)}{f_1(4)P(D)+f_2(4)P(N)}
$$



The equationm, simplified, now becomes the following:

$$
P(D|X=4)= \frac {.8(e^{-.5})}{(.8(e^{-.5}))(.2(e^{-.22})}=.752
$$

Thus we can claim the following :
The probability the company will issue a dividend given last year’s profit was 4% is approximately 75%

If you want, I can also show you the exact R code that reproduces this result or help you generalize the formula for any value of X.

## Stock Market Predictions Part 3 (Exercise 4.8.13)

Part A

```{r}
data("Weekly")
train <- Weekly$Year <= 2008
test  <- Weekly$Year > 2008
lda_fit <- lda(Direction ~ Lag2, data = Weekly, subset = train)
lda_pred <- predict(lda_fit, Weekly[test, ])$class
table(lda_pred, Weekly$Direction[test])
mean(lda_pred == Weekly$Direction[test])
```

Part B

```{r}
qda_fit <- qda(Direction ~ Lag2, data = Weekly, subset = train)

qda_pred <- predict(qda_fit, Weekly[test, ])$class

table(qda_pred, Weekly$Direction[test])

mean(qda_pred == Weekly$Direction[test])
```

Part C

With the provided accuracy measures, shown above, I would most likely suggest LDA, as it results in the highest accuracy, at 62%
